---
title: "linear_models"
author: "Angelica Bailey"
date: "2025-11-06"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```


Load the NYC airbnb data

```{r}
data(nyc_airbnb)
```

Look at the data/do some cleaning
```{r}
nyc_airbnb =
  nyc_airbnb |> 
  mutate(
    stars = review_scores_location / 2
  ) |> 
  rename(
    borough = neighbourhood_group
  ) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, room_type, neighbourhood)
```


do regression
```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```
R converts borough to factor(bronx, brooklyn, queens, manhattan)

do some additional cleaning

```{r}
nyc_airbnb =
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough), #make reference group the one that is most common, based on count
    room_type = fct_infreq(room_type)
  )

fit = lm(price ~ stars + borough, data = nyc_airbnb)
#changing interpretation of coefficients
```


Look at `lm` stuff
```{r, eval = FALSE}
summary(fit)
names(summary(fit))
summary(fit)[["coefficients"]]
summary(fit)[["df"]]

```


Look at cleaner `lm` stuff
```{r}
fit |> 
  broom::tidy() |> 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)
```

## Diagnostics

Look at residuals

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = resid)) +
  geom_histogram()


nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```


## Hypothesis testing

```{r}
fit |> 
  broom::tidy()
```

What about a categorical variable ?

```{r}
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)

fit_alt |> 
  broom::tidy()

anova(fit_null, fit_alt) |> 
  broom::tidy()
```


## Interactions vs nested data
```{r}
#lets allow stars effect to differe between borough and effect of room type to differ between borough
fit_interactions =
  lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

fit_interactions |> 
  broom::tidy()
```

Lets just look at Brooklyn first
```{r}
#just looking at the effect of stars in brooklyn
nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Bronx") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```


Write a short function

```{r}
lm_airbnb = function(df) {
  
  lm(price ~ stars + room_type, data = df)
  
}

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm_airbnb() |> 
  broom::tidy()
```


Create a list of dataframes and iterate to fit the model each time
```{r}
nested_lm_results = 
  nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    fits = map(data, lm_airbnb),
    results = map(fits, broom::tidy)
    ) |> 
  select(borough, results) |> 
  unnest(results)
```

Do some untidying
```{r}
nested_lm_results |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```



Use an *anonymous* function instead of `lm_airbnb`
```{r}
nested_lm_results = 
  nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(fits, broom::tidy)
    ) |> 
  select(borough, results) |> 
  unnest(results)

nested_lm_results |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )

```


Lets do an even more extreme examples
```{r}
#look neighborhood by neighborhood in manhattan
manhattan_analysis =
  nyc_airbnb |> 
  filter(
    borough == "Manhattan"
  ) |> 
  nest(data = -neighbourhood) |> 
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(fits, broom::tidy)
  ) |> 
  select(neighbourhood, results) |> 
  unnest(results)
```

Make a plot

```{r}
manhattan_analysis |> 
  filter(term == "stars") |> 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, estimate)
  ) |> 
  ggplot(aes(x = neighbourhood, y = estimate)) +
  geom_point()
```




